---
title: "GLP-1 Obesity Evidence Toolkit: Trial Design & Network Meta-Analysis"
subtitle: "Market positioning determines trial design"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7.5, fig.height = 4.2)
source("../R/sample_size.R")
source("../R/nma_helpers.R")
library(ggplot2)
library(DT)
```

## Why this exists

Parexel's *"Navigating to 2030"* playbook for obesity therapies opens with a
line worth taking literally: **decision-making must begin with precise market
positioning, which in turn determines optimal trial designs.** This toolkit
takes that claim and makes it computable. It does two things a sponsor or a
CRO's biostatistics team would actually need to do before writing a Phase 3
protocol in this drug class:

1. **Ask what sample size the primary efficacy comparison really requires** --
   and check that against the FDA's 2025 draft guidance safety-database floor,
   since obesity trials are unusually likely to be safety-, not efficacy-,
   constrained.
2. **Quantify where a candidate would land competitively** using a network
   meta-analysis of publicly reported Phase 3 results for the drugs already on
   the market or about to be, since head-to-head trials against every
   competitor are never feasible.

Everything below is computed live from the source data and code in this
repository, not pasted in from a report.

## Part A: how big does the trial actually need to be?

```{r ss-table}
tab <- scenario_table(deltas = c(3, 5, 8, 10, 12, 15, 18), sd = 10, power = 0.90,
                       alpha = 0.05, ratio = 2, dropout = 0.20)
knitr::kable(tab, digits = 0,
             col.names = c("Assumed delta (pp)", "SD", "Power", "Alpha", "Ratio", "Dropout",
                           "Efficacy N (control)", "Efficacy N (active)",
                           "Safety floor (control)", "Safety floor (active)",
                           "Final N (control)", "Final N (active)", "Final N (total)",
                           "Binding constraint"))
```

```{r ss-plot}
tab$eff_total <- tab$efficacy_driven_control + tab$efficacy_driven_active
ggplot(tab, aes(x = delta)) +
  geom_line(aes(y = eff_total, color = "Efficacy-driven total N"), linewidth = 1) +
  geom_hline(aes(yintercept = 4500, color = "FDA safety floor (total)"), linewidth = 1, linetype = "dashed") +
  scale_color_manual(values = c("Efficacy-driven total N" = "#3B6E8F", "FDA safety floor (total)" = "#C0392B")) +
  labs(x = "Assumed placebo-adjusted effect (percentage points)", y = "Total N", color = NULL,
       title = "Across the plausible effect-size range, the safety floor dominates") +
  theme_minimal(base_size = 12) + theme(legend.position = "top")
```

Across the entire range of effect sizes actually observed in this drug class
(roughly 8-20 placebo-adjusted percentage points, see Part B), the efficacy
comparison alone would be answerable with a few dozen patients per arm. What
actually sets the trial's size is the FDA's 2025 draft-guidance requirement of
&ge;3,000 subjects exposed to the investigational drug and &ge;1,500 to
placebo for &ge;1 year at the maintenance dose -- consistent with the observed
sizes of STEP 1 (n=1,961) and SURMOUNT-1 (n=2,539). A protocol synopsis that
sizes the trial from the power calculation alone, without checking it against
the safety floor, understates the true enrollment target.

## Part B: where does a candidate land competitively?

```{r nma-fit}
armdata <- read_armdata("../data/glp1_trials_armlevel.csv")
net <- build_network(armdata, reference = "Placebo")
fit <- fit_nma(net)
fd <- forest_data(fit)
sc <- sucra(fit)
```

```{r nma-forest}
fd$treatment <- factor(fd$treatment, levels = fd$treatment[order(fd$estimate)])
ggplot(fd, aes(x = estimate, y = treatment)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.15) +
  geom_point(size = 2.5) +
  labs(x = "% body weight change vs Placebo (pp, 95% CI)", y = NULL,
       title = "Network estimate of weight change at ~68-72 weeks") +
  theme_minimal(base_size = 12)
```

```{r nma-sucra-table}
knitr::kable(sc, digits = 3, col.names = c("Treatment", "SUCRA", "P(best)", "Mean rank"))
```

The network connects four published Phase 3 trials (STEP 1, STEP 8,
SURMOUNT-1, SURMOUNT-5) through a shared Placebo node and one direct
head-to-head link (SURMOUNT-5: tirzepatide vs. semaglutide), so the direct
tirzepatide-vs-semaglutide evidence can be checked against the indirect
route through Placebo. The global consistency test gives
Q = `r round(fit$Q, 2)` on `r fit$df` df (p = `r round(fit$Q_pvalue, 3)`),
i.e. no evidence against the transitivity assumption in this small network --
though with only one closed loop this is a weak check, not a guarantee (see
Limitations).

```{r nma-league}
DT::datatable(as.data.frame(league_table(fit)),
              caption = "League table: row minus column, 95% CI, percentage points",
              options = list(dom = "t", scrollX = TRUE))
```

## Limitations (read this before citing any number above)

- **Standard deviations are assumed, not sourced**, at 10 percentage points
  for every arm. Most public trial summaries (press releases, news coverage)
  report means and, at best, a treatment-difference confidence interval --
  not per-arm SDs. Where a difference and its CI were directly reported
  (e.g. STEP 1: -12.4pp, 95% CI -13.4 to -11.5), the implied SD is close to
  10, which is why that value was used throughout; this should be replaced
  with trial-reported SDs (or full summary statistics from the original
  papers/supplements) before this analysis is used for anything beyond
  illustration.
- **This is a fixed-effect model.** The project design document specifies a
  Bayesian random-effects NMA; this repository's from-scratch implementation
  (`R/nma_helpers.R`) reproduces the fixed-effect special case only, because
  `netmeta`/`gemtc` could not be installed from CRAN in this environment. See
  the commented block at the end of `analysis/02_nma_analysis.R` for how to
  swap in `netmeta` if you have full package access.
- **SURMOUNT-5's tirzepatide arm used the maximum tolerated dose (up to
  15 mg)**, which is treated here as equivalent to the SURMOUNT-1 15 mg arm.
  This is a simplification.
- **This is not a systematic review.** Four trials were chosen because they
  form a connected, illustrative network with a genuine head-to-head link,
  not because a systematic search was run. A real submission-grade NMA
  requires the full PICOS-driven search described in the project design
  document.

## Data sources

```{r sources}
knitr::kable(unique(armdata[, c("study", "year", "journal", "source_url")]),
             row.names = FALSE, col.names = c("Trial", "Year", "Journal", "Source"))
```
